taxadb: A High-Performance Local Taxonomic Database Interface

### 2020-02-12 paper/manuscript.Rmd
add acknowledgments and author contributions

---
title: "taxadb: A High-Performance Local Taxonomic Database Interface"
author:
  - name: "Kari Norman"
    affiliation: ucb
  - name: "Scott Chamberlain"
    affiliation: ropensci
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |
  A familiar and growing challenge in ecological and evolutionary research is that of reconciling scientific names of relevant taxa when combining data from separate sources. While this problem is already well understood and numerous naming authorities have been created to address the issue, most researchers lack a fast, consistent, and intuitive way to reconcile taxonomic names.  Here, we present `taxadb` R package to address this gap.  In contrast to existing tools, `taxadb` provides the following:  
  1) `taxadb` accesses established naming authorities to resolve synonyms, IDs, and hierarchical classification.  
  2) `taxadb` creates a local database, managed automatically from within R, that provides fast operations on millions of taxonomic names.  
  3) `taxadb` provides a consistent and intuitive data format.  
  4) `taxadb` is built on a simple, extensible and language agnostic design that can easily accommodate new authorities.

@@

>>>>

---
title: "taxadb: A High-Performance Local Taxonomic Database Interface"
author:
  - name: "Kari E A Norman"
    affiliation: ucb
  - name: "Scott Chamberlain"
    affiliation: ropensci
  - name: "Carl Boettiger"
    affiliation: ucb, ropensci
address:
  - code: ucb
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  - code: ropensci
    address: "The rOpenSci Project, University of California Berkeley, Berkeley CA 94720-3114, USA" 
abstract: |
  A familiar and growing challenge in ecological and evolutionary research is that of reconciling scientific names of relevant taxa when combining data from separate sources. While this problem is already well understood and numerous naming authorities have been created to address the issue, most researchers lack a fast, consistent, and intuitive way to reconcile taxonomic names.  Here, we present `taxadb` R package to address this gap.  In contrast to existing tools, `taxadb` provides the following:  
  1) `taxadb` accesses established naming authorities to resolve synonyms, IDs, and hierarchical classification.  
  2) `taxadb` creates a local database, managed automatically from within R, that provides fast operations on millions of taxonomic names.  
  3) `taxadb` provides a consistent and intuitive data format.  
  4) `taxadb` is built on a simple, extensible and language agnostic design that can easily accommodate new authorities.

@@

# Acknowledgments
We thank the many researchers who contributed to the data and infrastructure of the various taxonomic providers we access through our package. Support for the development of this package was provided by United States Department of Energy through the Computational Sciences Graduate Fellowship (DOE CSGF) under grant number DE-FG02-97ER25308 awarded to K.E.A.N..

# Authors' Contributions
K.E.A.N., S.C., and C.B. contributed to conceptual development of the package. K.E.A.N. and C.B. developed the package and contributed to the manuscript.
